---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CytOpT


<!-- badges: start -->
[![R-CMD-check](https://github.com/sistm/CytOpt-R/workflows/R-CMD-check/badge.svg)](https://github.com/sistm/CytOpt-R/actions)
<!-- badges: end -->

`CytOpT` uses regularized optimal transport to directly estimate the different cell population proportions from a biological sample characterized with flow cytometry measurements.

## Overview

`CytOpT` is an `R` package that provides a new algorithm relying regularized optimal transport to directly estimate the different cell population proportions from a biological sample characterized with flow cytometry measurements. Algorithm is based on the regularized Wasserstein metric to compare cytometry measurements from different samples, thus accounting for possible mis-alignment of a given cell population across sample (due to technical variability from the technology of measurements).

The main function of the package is `CytOpT()`.

The methods implemented in this package are detailed in the following article:

> Paul Freulon, Jérémie Bigot, Boris P. Hejblum. CytOpT: Optimal Transport with Domain Adaptation for Interpreting Flow Cytometry data
> https://arxiv.org/abs/2006.09003

## Installation

You can install the development version of CytOpT like so:
```{r GitHub-install, eval=FALSE}
remotes::install_github("sistm/CytOpT-R")
library(CytOpT)
```


## Example

This is a basic example of `CytOpt` usage:

### Data import
```{r example 1, message=FALSE}
library(CytOpT)
# Load source Data
X_source <- read.csv('tests/ressources/W2_1_values.csv')
Lab_source <- read.csv('tests/ressources/W2_1_clust.csv')[, 2]

# Load target Data
X_target <- read.csv('tests/ressources/W2_7_values.csv')
Lab_target <- read.csv('tests/ressources/W2_7_clust.csv')[, 2]
```

```{r example 2}
# Define the true proportions in the target data set X_source
theta_true <- rep(0,10)
for (k in 1:10) theta_true[k] <- sum(Lab_target == k)/length(Lab_target)
```

### Proportion estimations using *optimal transport* and *minmax swapping* procedures
```{r example 3, message=FALSE}
# Run CytOpt and compare the two optimization methods
res <- CytOpT(X_source, X_target, Lab_source, theta_true=theta_true, method='both')
```


```{r example res}
res$proportions
```


```{r example BA plot}
Bland_Atlman(res$proportions)
```


